---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbatext

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/MattCowgill/rbatext/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MattCowgill/rbatext/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This R package enables easy access to the full text of the [Reserve Bank of Australia's](https://www.rba.gov.au/) monetary policy [decisions](https://www.rba.gov.au/monetary-policy/int-rate-decisions/2025/) and [minutes](https://www.rba.gov.au/monetary-policy/rba-board-minutes/2025/).  

The package contains two key functions - `read_rba_decisions()` and `read_rba_minutes()`.

## Installation

You can install `rbatext` from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("MattCowgill/rbatext")
```

## Reading RBA decisions and minutes

The functions `read_rba_decisions()` and `read_rba_minutes()` do what they say on the tin:

```{r, eval = FALSE}
library(rbatext)
```
```{r, include = FALSE}
devtools::load_all()
```
```{r}
decisions <- read_rba_decisions()

decisions
```

This returns a tidy tibble containing the full text of each monetary policy decision since 1990. 

```{r}
minutes <- read_rba_minutes()

minutes
```

This returns a tidy tibble with the minutes of monetary policy meetings (first of the RBA Board, now of the RBA Monetary Policy Board) since 2006.

The package contains historical text stored as an internal data object. The functions, when run, scrape the RBA website for any new minutes/decisions.

## Example 1: Working with this text data

Here's an example of how to use the wonderful [`{tidytext}`](https://www.tidytextmining.com) package to work with this text data. 

First we 'unnest' the text statements, counting the number of occurrences of each word in each year:
```{r}
library(dplyr)
library(lubridate)
library(tidytext)

raw_word_counts <- decisions %>% 
  mutate(year = year(date)) %>% 
  unnest_tokens(word, text) %>% 
  group_by(year, word) %>% 
  count() 

raw_word_counts
```

Then we remove 'stop words' (common words like 'the' or 'if'), as well as removing numbers from the text:

```{r}
word_counts <- raw_word_counts %>% 
  anti_join(stop_words,
            by = "word") %>% 
  filter(is.na(suppressWarnings(as.numeric(word))))

word_counts

```

Then we calculate the [term frequency-inverse document frequency](https://www.tidytextmining.com/tfidf) of each word for each year. This is the number of times a word is used in a given year, adjusted by the number of times the word is used across all years.

```{r}

year_tfidfs <- word_counts %>% 
  bind_tf_idf(word, year, n)

year_tfidfs

```

Now we have this measure of which words best characterise the statements issued in a given year. Let's create a table showing the top 3 words for each year, ranked by tf-idf:

```{r}
year_tfidfs %>% 
  group_by(year) %>% 
  mutate(rank = rank(desc(tf_idf),
                     ties.method = "first")) %>% 
  filter(rank <= 5) %>% 
  arrange(rank) %>% 
  group_by(year) %>% 
  summarise(top_words = paste(word, collapse = ", ")) %>% 
  knitr::kable()
```


## The small print 
*Prior to December 2007, a media release was only issued if the Board decided to adjust monetary policy. Since that time, statements have been issued after each meeting, regardless of the outcome.

Note that this package has an 'experimental' badge. The package works and is stable. However, changes to the format or structure of the RBA website would break the package's key function. For this reason, the function is somewhat 'brittle'. I do not intend to submit this package to CRAN. I may at some point merge this package with [readrba](https://github.com/MattCowgill/readrba).
