---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rbatext

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/MattCowgill/rbatext/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MattCowgill/rbatext/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The Board of the Reserve Bank of Australia is responsible for monetary policy. It meets 11 times each year. After each meeting, the Governor of the RBA issues a statement outlining the reasons the Board did, or did not, adjust interest rates at that meeting.* 

This R package contains the text of those decisions in a tidy tibble. It also contains a function - `read_rba_decisions()` - that makes it easy to scrape future decisions from the RBA website. 


## Installation

You can install `rbatext` from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("MattCowgill/rbatext")
```

## RBA decisions

The key function in the package is `read_rba_decisions()`. You use it like this:

```{r}
library(rbatext)

decisions <- read_rba_decisions()

decisions
```

This returns a tidy tibble containing the full text of each monetary policy decision since 1990. 

### The small print 
*Prior to December 2007, a media release was only issued if the Board decided to adjust monetary policy. Since that time, statements have been issued after each meeting, regardless of the outcome.

Note that this package has an 'experimental' badge. The package works and is stable. However, changes to the format or structure of the RBA website would break the package's key function. For this reason, the function is somewhat 'brittle'. I do not intend to submit this package to CRAN. I may at some point merge this package with [readrba](https://github.com/MattCowgill/readrba).
